---
title: "Regression Models Course Project"
author: "Andrew Januszewski"
output: pdf_document
geometry: margin = 1cm
---
\fontsize{9}{3}

##**Setup**
```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(GGally)
knitr::opts_chunk$set(echo = TRUE)
data("mtcars")
```

##**Executive Summary**

    Moter Trend, a magazine about the automobile industry, is interested in examining the relationship between
    a set of variables and miles per gallon (MPG). They are particularly interested in answering the following
    questions:
        * Is an automatic or manual transmission better for MPG?
        * Can we quantify the difference between automatic and manual transmissions?
    This analysis concludeds manual transmissions are better for MPG than automatic. The three best predictors
    of MPG were transmission type, vehicle weight, and 1/4 mile time. Holding the other two variables constant,
    manual transmissions are ~2.94 more MPG efficient than automatics.
      
##**Exploratory Data Analysis**

    As shown in appendix figure one, transmission type (am, 0 = automatic, 1 = manual) and MPG share a 
    moderately strong, positive correlation. However. we also see other variables having similar or stronger
    correlation values with MPG (bottom row of figure one). As such, these predictors should be controlled
    for during any predictive model fitting.
```{r, eval = F}
ggcorr(mtcars, name = 'Correlation', label = T)
```
    
    For instance, consider the strong negative correlation between horsepower (hp) and MPG. It suggests as 
    horsepower increases, MPG decreases. Figure two in the appendix confirms this notion (black line). 
    However, notice that our prediction changes significantly when we include transmission type (am) as a 
    variable (blue - manual, red - automatic). Also, take note that the relationship between horsepower
    and MPG appears to be non-linear.
    
```{r, eval = F}
ggplot(mtcars, aes(y = mpg, x = hp, color = factor(am))) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ x, color = 'black', se = F) +  
  geom_smooth(method = 'lm', se = F) +
  labs(y = 'MPG', x = 'Horsepower') +
  theme(legend.position = 'none')
```
    
    We can further prove these observations by examining the residuals of each model. Appendix figure three 
    displays a reduction in Residual Standard Error (RSE) from 3.863 to 2.909. Additionally, explained error
    (R-squared) increased with the addition of am as a predictor. As we know, R-squared will continually 
    grow with every variable added. However, Adjusted R-squared accounts for that and increased from
    0.5892 to 0.767. Furthermore, both predictors are siginificant and a nested model confirms that in
    figure four.
    
```{r, eval = F}
fit_mpg_hp <- lm(data = mtcars, mpg ~ hp - 1)
fit_mpg_hp_am <- lm(data = mtcars, mpg ~ hp + factor(am) - 1)
summary(fit_mpg_hp)
summary(fit_mpg_hp_am)
anova(fit_mpg_hp, fit_mpg_hp_am)
```
    
    A plot of the residuals in appendix figure five also shows much less dispersion around zero with
    the addition of am as a predictor.
```{r, eval = F}
plot(fit_mpg_hp$residuals)
plot(fit_mpg_hp_am$residuals)
```
    
##**Data Modeling**

    We will fit and compare two linear models to answer the questions. The first will be extremely 
    simple while the second will leverage more variables choosen with some automobile knowledge. 
    
    For the simple model, we use transmission type as the sole predictor of MPG. It appears transmission type
    is significant, but this simple model only explains ~34% of variance.
```{r, echo = T}
fit_lm_simple <- lm(data = mtcars, mpg ~ factor(am))
summary(fit_lm_simple)
```
    
    For the second model, we will use some subject matter knowledge to include additional predictors. This 
    knowledge allows us to discount some combinations of highly correlated variables such as displacement 
    and cylinders as they are semi-interchangeable (displacement is the volume of all the pistons inside 
    the cylinders, more cylinders = higher displacement). Both are also highly correlated with horsepower. 
    Using this logic, we can exclude several other combinations. These could also be verified with the 
    Variance Inflation Factor (VIF), but we're running out of allotted pages for this assignment so we'll
    use the step function to find the most crucial variables.
    
    Again, transmission type is significant and the model now explains 83% of variance. This means 17% of
    unexplained variance could come from variables outside the model and dataset. Keeping this uncertainty
    in mind, we can conclude that manual transmissions are better for MPG. Holding the other two most 
    explanatory variables constant (wt and qsec), manual transmissions are ~2.94 more MPG efficient.
```{r, echo = T}
fit_lm_vars <- step(lm(data = mtcars, mpg ~ . - cyl - disp - gear - drat), trace = 0)
summary(fit_lm_vars)
```    

\pagebreak

##**Appendix**

###**Figure 1:**
```{r, echo = F, fig.width = 5, fig.height = 4.5}
ggcorr(mtcars, name = 'Correlation', label = T) +
  labs(subtitle = ' 0.0 - No Linear Relationship \n 0.3 - Weak \n 0.5 - Moderate \n 0.7 - Strong \n 1.0 - Exact') +
  theme(plot.subtitle = element_text(vjust = -40))
```

###**Figure 2:**
    Black = MPG predicted by horsepower
    Blue = MPG predicted by horsepower for manual transmission
    Red = MPG predicted by horsepower for automatic transmission
```{r, echo = F, fig.width = 6, fig.height = 4}
ggplot(mtcars, aes(y = mpg, x = hp, color = factor(am))) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ x, color = 'black', se = F) +  
  geom_smooth(method = 'lm', se = F) +
  labs(y = 'MPG', x = 'Horsepower') +
  theme(legend.position = 'none')
```

###**Figure 3:**
```{r, echo = F}
fit_mpg_hp <- lm(data = mtcars, mpg ~ hp)
fit_mpg_hp_am <- lm(data = mtcars, mpg ~ hp + factor(am))

summary(fit_mpg_hp)
summary(fit_mpg_hp_am)
```

###**Figure 4:**
```{r, echo = F}
anova(fit_mpg_hp, fit_mpg_hp_am)
```

###**Figure 5:**
```{r, echo = F}
plot(fit_mpg_hp$residuals)
```

```{r, echo = F}
plot(fit_mpg_hp_am$residuals)
```